---
title: "Optimize denovo STACKS pipeline"
output: rmarkdown::html_vignette
fig_width: 6 
fig_height: 4
vignette: >
  %\VignetteIndexEntry{Optimize denovo STACKS pipeline}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(dev = "png")
```

```{r}
#load packages
library(RADstackshelpR)
library(gridExtra)
```

##Step 1: Demultiplex each sample using 'process_radtags'.
##Step 2: Samples with a demultiplexed file size of < 2 megabytes simply don't have enough information to be relevant downstream, and can be removed right away. (Note: this will help ensure that there are enough loci retained at an 80% completeness cutoff to make informed optimization decisions)
##Step 3: Iterate over values of m ranging from 3-7, while leaving all other parameters at default values.
##Step 4: Visualize the output of these 5 runs and determine the optimal value for m.
```{r, dpi=300, out.width="100%",out.height="100%", fig.align = "center"}
#optimize_m function will generate summary stats on your 5 iterative runs
#input can be full path to each file or just the file name if your working directory contains the files
m.out<-optimize_m(m3="/Users/devder/Desktop/hipposideros/m_3.vcf",
           m4="/Users/devder/Desktop/hipposideros/m_4.vcf",
           m5="/Users/devder/Desktop/hipposideros/m_5.vcf",
           m6="/Users/devder/Desktop/hipposideros/m_6.vcf",
           m7="/Users/devder/Desktop/hipposideros/m_7.vcf")

#visualize the effect of varying m on the depth of each sample
vis_depth(output = m.out)
#visualize the effect of varying m on the number of SNPs retained
vis_snps(output = m.out, stacks_param = "m")

#visualize the effect of varying m on the number of loci retained
vis_loci(output = m.out, stacks_param = "m")

#3 is the optimal m value, and will be used next to optimize M
```

##Step 5: Iterate over values of M ranging from 1-8, setting m to the optimal value (here 3).
##Step 6: Visualize the output of these 8 runs and determine the optimal value for M.
```{r, dpi=300, out.width="100%",out.height="100%", fig.align = "center"}
#optimize M
M.out<-optimize_bigM(M1="/Users/devder/Desktop/hipposideros/M1.vcf",
           M2="/Users/devder/Desktop/hipposideros/M2.vcf",
           M3="/Users/devder/Desktop/hipposideros/M3.vcf",
           M4="/Users/devder/Desktop/hipposideros/M4.vcf",
           M5="/Users/devder/Desktop/hipposideros/M5.vcf",
           M6="/Users/devder/Desktop/hipposideros/M6.vcf",
           M7="/Users/devder/Desktop/hipposideros/M7.vcf",
           M8="/Users/devder/Desktop/hipposideros/M8.vcf")

#visualize the effect of varying M on the number of SNPs retained
vis_snps(output = M.out, stacks_param = "M")

#visualize the effect of varying M on the number of polymorphic loci retained
vis_loci(output = M.out, stacks_param = "M")
```

##Step 7: Iterate over values of n ranging from M-1, M, M+1, setting m and M to the optimal values (here 3 and 2, respectively).
##Step 8: Visualize the output of these 3 runs and determine the optimal value for n.
```{r, dpi=300, out.width="100%",out.height="100%", fig.align = "center"}
#optimize n
n.out<-optimize_n(nequalsMminus1="/Users/devder/Desktop/hipposideros/n1.vcf",
           nequalsM="/Users/devder/Desktop/hipposideros/n2.vcf",
           nequalsMplus1="/Users/devder/Desktop/hipposideros/n3.vcf")

#visualize the effect of varying n on the number of SNPs retained
vis_snps(output = n.out, stacks_param = "n")

#visualize the effect of varying n on the number of polymorphic loci retained
vis_loci(output = n.out, stacks_param = "n")

```

#Finally, make a single figure showing the optimization process of all three parameters
```{r, dpi=300, out.width="100%",out.height="100%", fig.align = "center"}
gl<-list()
gl[[1]]<-vis_depth(output = m.out)
gl[[2]]<-vis_snps(output = m.out, stacks_param = "m")
gl[[3]]<-vis_loci(output = m.out, stacks_param = "m")
gl[[4]]<-vis_snps(output = M.out, stacks_param = "M")
gl[[5]]<-vis_loci(output = M.out, stacks_param = "M")
gl[[6]]<-vis_snps(output = n.out, stacks_param = "n")
gl[[7]]<-vis_loci(output = n.out, stacks_param = "n")

grid.arrange(
  grobs = gl,
  widths = c(1,1,1,1,1,1),
  layout_matrix = rbind(c(1,1,2,2,3,3),
                        c(4,4,4,5,5,5),
                        c(6,6,6,7,7,7))
)

```


